---
title: "HW_1"
author: "Stephen Skasko"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

#1
```{r}
# a)

a = 1/1i
a

# b)

b = exp(1i * pi)
b

# c)
c = cos(pi *(1:4)/2)
c

```
\newpage

#2

The warning is from trying to add two different vectors with different dimensions 
The error tells us that it may give us unexpected results because of the length difference
```{r}
output = c(8,3,2) + c(2,4)
output

```
\newpage

#3
```{r}
v1 = rep(c(1,2), e = 3)
v2 = rep(c(1,2), t = 3)

v1
v2
```
\newpage


#4

This tells us to take remainder of each number from the sequence of 1 to 20 values / 10,and %% 1 returns the decimal of the number 

```{r}
output = ( 1 : 20 / 10 ) %% 1
output

```
\newpage

#5

It seems that gas production in Canada had a significant increase from 2002 to 2006. The two biggest outliers are from 1996-2000 with a huge drop in 2000 and the other one is from 2002-2006 with a huge increase of volume. Based on the graph, we are able to find a slight trend potentially from 1995-2005 and then repeat from significantly low values to high throughout every few years. With these factors, we can understand that this graph is cyclicity. 
```{r}
library(fpp3)

#a)
data = data("canadian_gas")

# b)
extract = ts(canadian_gas$Volume, start = c(1960, 1), f = 12) #12 months in year

# c)
subset = ts(extract, start = c(1994, 1), end = c(2004, 12))

# d)
plot(subset, main = "Canadian Monthly Gas Production from 1994 to 2004", ylab = "Volume")

```
\newpage

#6

Graph one for lynx: 
In the graph, we can see a trend happening with the number of lynx every approx 40 years with no real outliers as the trend seems more consistent. The graph strikes me as cyclicity and potentially seasonal since it seems to fluctuate higher when hitting certain months in the spring or early summer when following along a 12 month period for number of lynx. 

Graph two for sunspotz: 
Although not perfect, there is a slight potential trend until there are a huge few outlier dips around 100-150 years. Additionally, there seems to be an outlier at approx 200 around 420 years. With this, it looks like there is only a cyclicity trend going on. 

```{r}
# a)
library(fma)
library(astsa)
help('lynx')
help('sunspotz') 

# b)
ts_lynx = ts(lynx)
ts_sunspot = ts(sunspotz)

plot(ts_lynx, main = "Canadian Lynx Trapping from 1821 to 1934", ylab = "Number of Lynx")

plot(ts_sunspot, main = "Yearly Sunspot Counts from 1700 to 1988", ylab = "Number of Sunspots")

```
\newpage

#7

a) In the graph, there is a slight trend, but becomes slightly off around 50-70 in years and returns to the trend, where an outlier may include at 60 years. The moving average does seem to showcase a slight trend as well, making the noise slightly smooth and following the pattern of the trend. 

b) In the graph, there is a trend that fluctuates up and down from -1 to +1 over time. Since the graph and series is consistent, there seems to be no outlier. Additionally, there would be no seasonal trend but only cyclicity tends. Since the moving average is completely straight, meaning its a very smooth and is less noisy. 

c) In the graph, there seems to be no trends at all, with heavy outliers at +4 and -4. The moving average is very noisy meaning it can be hard to detect a trend and is not smooth as well for the graph.

d) Based on the graphs, we can see that they are all different in their own ways, where a) has a slight trend, b) has a steady trend, and c) has no trends. The outliers in all three of these are also similar to the trends, where a) shows a potential for an outlier, b) shows no outliers based on the steadiness of the graph, and c) shows many different points but with a few noticeable outliers. Lastly, each graphs noise levels show different results, where a) shows a slightly noisy and smooth line, b) shows a perfectly stable line with no noise, and c) shows an unstable and noisy moving average line. 

```{r}
set.seed(1000)


n = 100

wt = rnorm(n)

xt_a = numeric(n)
for (i in 3:n) {
  xt_a[i] = -0.9 * xt_a[i-2] + wt[i]
}

xt_b = cos(2 * pi * (1:n) / 4)
xt_c = xt_b + wt


va_xt_a = stats::filter(xt_a, rep(1/4, 4), sides = 1)
va_xt_b = stats::filter(xt_b, rep(1/4, 4), sides = 1)
va_xt_c = stats::filter(xt_c, rep(1/4, 4), sides = 1)


# 1:n helps from one x-axis to n number of x-values (100), while type = "l" is for lines, if not, all you get are points 
plot(1:n, xt_a, type = "l", xlab = "Time", ylab = "Xt")
lines(1:n, va_xt_a, col = "red", lty = 2)

plot(1:n, xt_b, type = "l", xlab = "Time", ylab = "Xt")
lines(1:n, va_xt_b, col = "red", lty = 2)

plot(1:n, xt_c, type = "l", xlab = "Time", ylab = "Xt")
lines(1:n, va_xt_c, col = "red", lty = 2)



```





